Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-03-15 ltx/BrainIDE.tex
Explaination to the derivation
added/updated discrete space-time

The average firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
f\left( {{v_t}\left( {r'} \right)} \right) = \frac{{{\nu _0}}}{{1 + \exp \left( {\beta \left( {{v_0}\left( r \right) - {v}\left(r',\bar{t} \right)} \right)} \right)}}.
\end{equation}
%\begin{equation}\label{RateOfPreSynPops}
%p\left( {v\left( {r,\bar t} \right)} \right) = {\nu_0}u\left( {v\left( {r,\bar t} \right) - {v_0}\left( r \right)} \right)
%\end{equation}
%\begin{equation}\label{HeavisideDef}
%u\left( {v\left( {r,\bar t} \right) - {v_0}\left( r \right)} \right) = \left\{ {\begin{array}{*{20}{c}}
%   1 & {v\left( {r,\bar t} \right) \ge {v_0}\left( r \right)}  \\
%   0 & {otherwise}  \\
%\end{array}} \right.
%\end{equation}
Even though we will model  $v_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality when constructing the state-space representation in the next section. The parameters in equation~\ref{ActivationFunction} are detailed in Table 1. By combining equations~\ref{SpikesToPotential} and \ref{RateBasedInteractions} we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
v\left(r,t\right) = \int\limits_{-\infty}^t\int\limits_\Omega  h\left(t - t'\right)w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' dt'+e\left(r,t\right)
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
Lh\left( t \right) = \delta \left( t \right),
\end{equation}
where $L$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function. This provides the most general form
\begin{equation}\label{GenForm}
L\left( \frac{\partial }{dt} \right)v\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  + de\left( r,t \right),
\end{equation}
where $L$ is a polynomial of order $n$ with constant coefficients that provides an $n^{th}$ order model. Most studies considering this model use first or second order time derivatives. We shall use the first order model giving
\begin{equation}\label{FinalForm1}
dv\left( {r,t} \right) + \alpha v\left( r,t \right)dt = \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',\bar t \right)} \right)dr'dt}  +de\left( r,t \right).
\end{equation}
The PDE form of the model is
\begin{equation}\label{PDE}
\frac{\partial^2}{\partial t \partial r'} v\left( r,t \right) + \alpha \frac{\partial}{\partial r'} v\left( r,t \right) = w\left( r,r' \right)f\left( v\left( r',\bar t \right) \right)  + \frac{\partial}{\partial r'}de\left( {r,t} \right).
\end{equation}
The spatial derivative of the Weiner process can be calculated since it is smooth in space (colored) although it is temporally white. 
% \subsection{Multi-Layer Model}
% \begin{equation}\label{ExcitatoryLayer}
% \frac{dv_e\left( {r,t} \right)}{{dt}} + \alpha_ev_e\left( {r,t} \right) = \int_\Omega  {w_{ee}\left( {r,r'} \right)f\left( v_e\left( r',\bar t \right) \right) + w_{ei}\left( {r,r'} \right)f\left( v_i\left( r',\bar t \right) \right)dr'}  +\varepsilon_e\left( {r,t} \right)
% \end{equation}
% \begin{equation}\label{InhibitoryLayer}
% \frac{dv_i\left( {r,t} \right)}{{dt}} + \alpha_iv_i\left( {r,t} \right) = \int_\Omega  {w_{ie}\left( {r,r'} \right)f\left( v_e\left( r',\bar t \right) \right) + w_{ii}\left( {r,r'} \right)f\left( v_i\left( r',\bar t \right) \right)dr'}  +\varepsilon_i\left( {r,t} \right)
% \end{equation}
% The subscript  denotes connectivity from an inhibitory neural element to an excitatory neural element. We can write the model in matrix form
% \begin{equation}\label{matrixEquation}
% \left[ {\begin{array}{*{20}{c}}
%    {\frac{{d{v_e}\left( t \right)}}
% {{dt}}}  \
%    {\frac{{d{v_i}\left( t \right)}}
% {{dt}}}  \
% 
%  \end{array} } \right] = \int_\Omega  {\left[ {\begin{array}{*{20}{c}}
%    {{w_{ee}}\left( {r,r'} \right)} & {{w_{ei}}\left( {r,r'} \right)}  \
%    {{w_{ie}}\left( {r,r'} \right)} & {{w_{ii}}\left( {r,r'} \right)}  \
% 
%  \end{array} } \right]} \left[ {\begin{array}{*{20}{c}}
%    {f\left( {{v_e}\left( {r',\bar{t}} \right)} \right)}  \
%    {f\left( {{v_i}\left( {r',\bar{t}} \right)} \right)}  \
% 
%  \end{array} } \right]dr' - \left[ {\begin{array}{*{20}{c}}
%    {{\alpha _e}{v_e}\left( {r,t} \right)}  \
%    {{\alpha _i}{v_i}\left( {r,t} \right)}  \
% 
%  \end{array} } \right] + \left[ {\begin{array}{*{20}{c}}
%    {{\varepsilon _e}\left( t \right)}  \
%    {{\varepsilon _i}\left( t \right)}  \
% 
%  \end{array} } \right]
% \end{equation}
% 
% \begin{equation}\label{matrix_equation_in_vector_notation}
%    \frac{d\mathbf{V}(r,t)}{dt} = \int_\Omega \mathbf{W}(r,r')f\left(\mathbf{V}(r,t)\right)dr' - \mathbf{A}\mathbf{V}(r,t) + \mathbf{E}(r,t)
% \end{equation}
% 
% The iEEG observation equation is
% \begin{equation}\label{iEEGObservationEquation}
%     y(k,t) = \int_{\Omega} m(k,r)\left(v_e(r,t) + v_i(r,t)\right)dr' + w(k,t)
% \end{equation}{}

% \subsection{Alternate Model Description I}
% An alternate and equivalent way to remove the temporal integration using the Laplace transform . We start by considering a discrete neural mass $i$ with inputs with zero delay from another discrete mass $j$ and
% \begin{equation}\label{TwoDiscreteMasses}
% {v_i}\left( t \right) = \int_{ - \infty }^t {{h_j}(t - t'){w_{ij}}p\left( {{v_j}\left( t \right)} \right)dt'},
% \end{equation}
% where $w_{ij}$ can be considered the synaptic efficiency from $j$ to $i$. Now we take the Laplace transform $\mathcal{L}(v_i(t))=V(s)$ yielding
% \begin{eqnarray}
% % \nonumber to remove numbering (before each equation)
%   {V_i}\left( s \right) &=& {H_j}\left( s \right){w_{ij}}P\left( {{V_j}\left( s \right)} \right) \nonumber \\
%    &=& {\left( {s + \alpha} \right)^{ - 1}}{w_{ij}}P\left( {{V_j}\left( s \right)} \right)  \\
%   \Rightarrow s{V_i}\left( s \right) + \alpha{V_i}\left( s \right) &=& {w_{ij}}P\left( {{V_j}\left( s \right)} \right).
% \end{eqnarray}
% Now taking the inverse Laplace transform we get
% \begin{equation}\label{InverseLaplace}
% \frac{{d{v_i}\left( t \right)}}{{dt}} + \alpha{v_i}\left( t \right) = {w_{ij}}p\left( {{v_j}\left( t \right)} \right).
% \end{equation}
% Next we convert the model from discrete neural masses to a continuum by
% \begin{equation}\label{Discrete2Continuum}
% \frac{{dv\left( {r,t} \right)}}{{dt}} + \alpha{v\left( {r,t} \right)} = \int_\Omega  {w\left( {r,r'} \right)p\left( {v\left( {r',\bar t} \right)} \right)dr'}  + e\left( {s,t} \right).
% \end{equation}

%\subsection{Alternate Model Description II}
%\begin{figure*}
%\centering
%\includegraphics[scale=0.4]{NeuralMassFigure.eps}
%\caption[NeuralMassModel]{This figure describes the basic interactions of neural masses that constitute the neural field model.}
%\label{NeuralMassModel}
%\end{figure*}
%An alternative description of the model is as follows. The model can be thought of as a continuum of neural masses. To illustrate this idea, we shall first describe the interaction of discrete neural masses as seen in Figure~\ref{NeuralMassModel}. Assuming that post-synaptic potentials sum linearly, the membrane potential at the soma of neuron $i$ is described as
%\begin{equation}\label{VoltageFromPSPs}
%{v_i}\left( t \right) = \sum\nolimits_{j,k} {{w_{ij}}{h_{j}}\left( t - {t_k}\right)} u\left( {t - {t_k}} \right),
%\end{equation}
%where $w_{ij}$ is the synaptic weight from neuron $j$ to $i$, $h(t)$ is the synaptic response kernel, $S$ is the Heaviside step function and action potential arrive at neuron $i$ at times $t_k$. Alternatively the membrane potential can described by the convolution of synaptic response kernel with the pre-synaptic firing rate $r_j(t)$ where
%\begin{equation}\label{VoltageFromRate}
%{v_i}\left( t \right) = \sum\nolimits_j {\int_{ - \infty }^t {{w_{ij}}{h_j}\left( {t - t'} \right){r_j}\left( t \right)} } dt'.
%\end{equation}
%To convert the equation system into an integral differential system, we shall take derivative of~\ref{EqVoltageFromPSPs} yielding
%\begin{eqnarray}
%% \nonumber to remove numbering (before each equation)
%  \frac{{d{v_i}\left( t \right)}}{{dt}} &=& \sum\nolimits_{j,k} {\left( {{w_{ij}}{h_j}\left( t- {t_k} \right)\frac{d}{{dt}}u\left( {t - {t_k}} \right) + {w_{ij}}u\left( {t - {t_k}} \right)\frac{d}{{dt}}{h_j}\left( t- {t_k} \right)} \right)} \nonumber \\
%   &=& \sum\nolimits_{j,k} {{w_{ij}}{h_j}\left( t- {t_k} \right)\delta \left( {t - {t_k}} \right)}  - \sum\nolimits_{j,k} {{w_{ij}}u\left( {t - {t_k}} \right)\alpha_j{h_j}\left( t- {t_k} \right)}   \nonumber \\
%   &=& \sum\nolimits_{j,k} {{w_{ij}}\delta \left( {t - {t_k}} \right)}  - {v_i}\left( t \right)\sum\nolimits_j {\alpha_j} \nonumber  \\
%   &=& \sum\nolimits_j {{w_{ij}}{\nu_j}\left( t \right)}  - {v_i}\left( t \right)\sum\nolimits_j {\alpha_j} \nonumber \\
%   &=& \sum\nolimits_j {{w_{ij}}p\left( {{v_j}\left( t \right)} \right)}  - {v_i}\left( t \right)\sum\nolimits_j {\alpha_j}.
%\end{eqnarray}
%This form can easily be extended to represent a continuous field.

%\subsection{Assumptions and Simplifications}
%We assume that the propagation delay of action potentials is zero. This is a common assumption that is made when modeling intracortical circuits (\textcolor[rgb]{1.00,0.00,0.00}{refs}). As stated in Wilson and Cohen's original works, the time constant $\tau$ is much greater than the axonal propagation delays, therefore it is reasonable to approximate the the delays as being negligible. Recent studies have demonstrated the importance of having a fast, but finite, propagation velocity for models to make realistic and complete predictions of cortical dynamics~\cite{Atay2005}. The purpose of this paper, however, is not to precisely simulate complex cortical dynamics, but to create a useful platform where estimation of hidden states is possible for the purpose of detecting abnormal neural states with intentions of generating informed patient specific control sequences. Therefore, the model simply has to be `good enough' for this purpose. Starting with this simplified version of the general model is a logical choice. The next assumption we will make is spatial homogeneity such that synaptic interaction between neural populations is only dependant on their spatial separation, not on their spatial location. Again, this is a common assumption when dealing with intracortical connectivity (\textcolor[rgb]{1.00,0.00,0.00}{ref}). This assumption is not essential for our IDE estimation techniques, where a heterogeneous structure is possible. The IDE estimation is derived for a general heterogeneous structure however, for clarity of results we shall only present the heterogeneous case. We will assume the synaptic response kernel has the same shape (scaled by the connectivity kernel) irrespective of the respective presynaptic population. In this case, inhibition can be modeled by having a negative response kernel and excitation with a positive response kernel, where the connectivity kernel provides the (scaled) polarity. This reduces the model to a single field. We will also assume the activation function $p(.)$ is a Heaviside step function. This is in agreement with the work of Schiff and Sauer~\cite{schiff2008kalman}. With these assumptions, the model has the form
%\begin{equation}\label{SimplifiedModel}
%\frac{{dv\left( {r,t} \right)}}{{dt}} = \int_\Omega  {w\left( {r-r'} \right)f\left( {v\left( {r',t} \right)} \right)dr'}  - \alpha{v\left( {r,t} \right)} + e\left( {r,t} \right).
%\end{equation}
%In order to make the problem trackable, in terms of model order, we assume that we can decompose the field (membranes voltages) into a set of $n_\phi$ Gaussian basis functions. We will also assume that we know the location and width of the basis function that is required to decompose the field. The state of the system is the height basis functions, rather than the membrane potential. Therefore, the number of basis functions required to decompose the field is equal to the number of states. Also, we assume that we can represent the connectivity by a regular grid of three connectivity Gaussian basis functions (corresponding to short, mid and long range connectivity) of known width. The height of the connectivity basis functions ($\theta$ parameter) is to be estimated.
%\subsection{IDE Formulation}

%For the state-space derivation, to create an estimator, the model can be reformulated like so
%\begin{equation}\label{new_spatiotemporal_kernels}
%\frac{{dv\left( {r,t} \right)}}{{dt}} = \int_\Omega  {k\left( {r - r'} \right)\tilde v\left( {r',t} \right) - \alpha \delta \left( {r - r'} \right)\bar v\left( {r,t} \right)dr'}  + e\left( {r,t} \right)
%\end{equation}
%\begin{equation}\label{define_v_tilde}
%\tilde v\left( {r,t} \right) = \left\{ {\begin{array}{*{20}{c}}
%   {v\left( {r,t} \right)} & {v\left( {r,t} \right) \ge {v_0}}  \\
%   0 & {otherwise}  \\
%\end{array}} \right.
%\end{equation}
%\begin{equation}\label{define_v_bar}
%\bar v\left( {r,t} \right) = \left\{ {\begin{array}{*{20}{c}}
%   {v\left( {r,t} \right)} & {v\left( {r,t} \right) < {v_0}}  \\
%   0 & {otherwise}  \\
%\end{array}} \right.
%\end%equation}
% \begin{figure*}
% \centering
% \includegraphics[scale=0.3]{ExampleFieldDecomposition.eps}
% \caption[ExampleFieldDecomposition]{Illustration of a one dimensional field decomposition using 5 Gaussian basis functions, each weighted an element of the state vector. Each basis function has a width of 0.01 and is uniformly positioned with a separation of 0.25 (with arbitrary units).}
% \label{ExampleFieldDecomposition}
% \end{figure*}

%\section{State-Space Neural Field Model}
%In this section we shall reconstruct the model in a state-space form that is a novel representation of the neural field equations. The benefit of restating the model in this format is that it becomes easier to simulate and allows the connectivity to be inferred from data. \begin{equation}
% v_{t+1}(\mathbf r)=\int_{\Omega} \alpha \delta(\mathbf r-\mathbf r')\bar v_t(\mathbf r)d \mathbf r'+\int_{\Omega} w(\mathbf r -\mathbf r')\tilde v_t(\mathbf r')d\mathbf r'+e_t(\mathbf r)
%\label{eq:cortexfield}
%\end{equation}
%decomposing the field and the kernel we have
%\begin{equation}
%\bar v_t(\mathbf r)=\boldsymbol \phi^{\top}(\mathbf r)\bar{\mathbf x}_t
%\label{eq:cortexfieldlv0}
%\end{equation}
%\begin{equation}
%\tilde v_t(\mathbf r)=\boldsymbol \phi^{\top}(\mathbf r)\tilde{\mathbf x}_t=\nu_0
%\label{eq:cortexfieldgv0}
%\end{equation}
%\begin{equation}
% w(\mathbf r -\mathbf r')=\boldsymbol \theta^{\top}\psi(\mathbf r -\mathbf r')
%\label{eq:kerneldecomposition}
%\end{equation}
%Substituting (\ref{eq:cortexfieldlv0}), (\ref{eq:cortexfieldgv0}) and (\ref{eq:kerneldecomposition}) in (\ref{eq:cortexfield}) we get
%\begin{equation}
% \boldsymbol \phi^{\top}(\mathbf r)\bar{\mathbf x}_{t+1}+\boldsymbol \phi^{\top}(\mathbf r)\tilde{\mathbf x}_{t+1}=
%\int_\Omega\alpha\delta(\mathbf r- \mathbf r')\boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_td\mathbf r'+\int_\Omega\boldsymbol \theta^{\top}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')\tilde{\mathbf x}_td\mathbf r'+e_t(\mathbf r)
%\label{eq:decomposedcortexfield}
%\end{equation}
%\begin{equation}
% \boldsymbol \phi^{\top}(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha \boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_{t}+\boldsymbol \theta^{\top}\int_{\Omega}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')\tilde{\mathbf x}_td\mathbf r'+e_t(\mathbf r)
%\end{equation}
%by defining
%\begin{equation}
% \boldsymbol\Phi(\mathbf r)=\int_{\Omega}\boldsymbol\psi(\mathbf r -\mathbf r')\boldsymbol\phi^{\top}(\mathbf r')d\mathbf r'
%\label{eq:convolution}
%\end{equation}
%we have
%\begin{equation}
% \boldsymbol \phi^{\top}(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha \boldsymbol\phi^{\top}(\mathbf r)\bar{\mathbf x}_{t}+\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)\tilde{\mathbf x}_t+e_t(\mathbf r)
%\label{eq:cortexfieldwithPhi}
%\end{equation}
%Premultiplying (\ref{eq:cortexfieldwithPhi}) by $\boldsymbol\phi(\mathbf r) $, and integrating over $\mathbf r$
%\begin{equation}
%\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \phi^{\top}(\mathbf r)d\mathbf s(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \phi^{\top}(\mathbf r)d\mathbf r \bar{\mathbf x}_{t}+\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)d\mathbf r\tilde{\mathbf x}_t+\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r
%\label{eq:cortexfieldversion2}
%\end{equation}
%Now we define
%\begin{equation}
% \boldsymbol\Psi_x(\mathbf r)=\int_{\Omega}\boldsymbol\phi(\mathbf r)\boldsymbol\phi^{\top}(\mathbf r)d\mathbf r
%\label{eq:psix}
%\end{equation}
%\begin{equation}
%  \boldsymbol\Psi_{\theta}(\mathbf r)=\int_{\Omega} \boldsymbol\phi(\mathbf r)\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf r)d\mathbf r
%\label{eq:psitheta}
%\end{equation}
%substituting (\ref{eq:psix}) and (\ref{eq:psitheta}) in (\ref{eq:cortexfieldversion2}) we get
%\begin{equation}
%\boldsymbol\Psi_x(\mathbf r)(\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1})=\alpha\boldsymbol\Psi_x(\mathbf r) \bar{\mathbf x}_{t}+\boldsymbol\Psi_{\theta}(\mathbf r)\tilde{\mathbf x}_t+\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r
%\end{equation}
%pre-multiplying both sides by $\boldsymbol\Psi_x^{-1}$
%\begin{equation}
%\bar{\mathbf x}_{t+1}+\tilde{\mathbf x}_{t+1}=\alpha\bar{\mathbf x}_{t}+\boldsymbol\Psi_x^{-1}\boldsymbol\Psi_{\theta}\tilde{\mathbf x}_t+\boldsymbol\Psi_x^{-1}\int_{\Omega} \boldsymbol\phi(\mathbf r)e_t(\mathbf r)d\mathbf r
%\end{equation}
%????????????????????????????????
%\subsection{Observation equation}
%\begin{eqnarray}
% \mathbf y_t(\mathbf r)&=&\\
%&=&\int_{\Omega}H(\mathbf r-\mathbf r')v_t(\mathbf r')d\mathbf r'+\boldsymbol\epsilon_t \\
%&=&\int_{\Omega}H(\mathbf r-\mathbf r')\bar v_t(\mathbf r')d\mathbf r'+\int_{\Omega}H(\mathbf r-\mathbf r')\tilde v_t(\mathbf r')d\mathbf r'+\epsilon_t
%\label{eq:observationequation}
%\end{eqnarray}
%
% we decompose $ H(\mathbf r-\mathbf r')$ as
% \begin{equation}
%  H(\mathbf r-\mathbf r')= \mathbf c^{\top}\mu(\mathbf r-\mathbf r')
%\label{eq:decomposedobservationkernel}
%\end{equation}
%and substituting  (\ref{eq:cortexfieldlv0}), (\ref{eq:cortexfieldgv0}) and (\ref{eq:decomposedobservationkernel}) in (\ref{eq:observationequation}) we get
%\begin{equation}

>>>>

The average firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
f\left( {{v_t}\left( {r'} \right)} \right) = \frac{{{\nu _0}}}{{1 + \exp \left( {\beta \left( {{v_0}\left( r \right) - {v}\left(r',\bar{t} \right)} \right)} \right)}}.
\end{equation}

Even though we will model  $v_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality when constructing the state-space representation in the next section. The parameters in equation~\ref{ActivationFunction} are detailed in Table 1. By combining equations~\ref{SpikesToPotential} and \ref{RateBasedInteractions} we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
v\left(r,t\right) = \int\limits_{-\infty}^t\int\limits_\Omega  h\left(t - t'\right)w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' dt'+e\left(r,t\right)
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
Lh\left( t \right) = \delta \left( t \right),
\end{equation}
where $L$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function. This provides the most general form
\begin{equation}\label{GenForm}
L\left( \frac{\partial }{dt} \right)v\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  + de\left( r,t \right),
\end{equation}
where $L$ is a polynomial of order $n$ with constant coefficients that provides an $n^{th}$ order model. Most studies considering this model use first or second order time derivatives. We shall use the first order model giving
\begin{equation}\label{FinalForm1}
dv\left( {r,t} \right) + \alpha v\left( r,t \right)dt = \alpha\int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',\bar t \right)} \right)dr'dt}  +de\left( r,t \right).
\end{equation}

\section{Discrete Time and Space Model}
We begin the derivation of the state-space model with the continuous time, continuous space IDE neural field model. 
Spatio-temporal interactions can be described by the continuous space continuous time stochastic differential equation
\begin{equation}\label{differential eq}	
dv\left( r,t \right) = \alpha \left( -v\left( r,t \right) + \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t \right)} \right)dr'} \right) dt  + d e\left( r,t \right).
\end{equation}
To simplify the notation, but without loss of generality, \ref{FinalForm1} can be rewritten as
\begin{equation}\label{differential simplified}
	dv\left( r,t \right) = \alpha q\left(v(r,t) \right) dt  + d e\left( r,t \right)
\end{equation}
A one-step Euler method was used to approximate the nonlinear stochastic IDE to a discrete time form. 
\begin{equation}\label{euler}
	v\left(r,t\right) =  v(r,t_0) + \alpha \int_{t_0}^{t} q\left(v(r,\tau) \right) d\tau  + \int_{t_0}^{t} d e\left(r,\tau\right)
\end{equation}
Using an Euler approximation,~\ref{euler} can be approximated at each sampling time by
\begin{equation}\label{euler approx 1}
	v\left(r,t_{k+1} \right) =  v(r,t_k) + \alpha q\left(v(r,t_k)\right) (t_{k+1} - t_k) + (e(r,t_{k+1}) - e(r,t_k))
\end{equation}
which can be rewritten as
\begin{equation}\label{euler approx 2}
	v\left(r,t_k + \delta \right) =  v(r,t_k) + \delta \alpha q\left(v(r,t_k)\right) + e'(r,t_k)
\end{equation}
where $e'(r,t_k)$ is a zero mean Gaussian random variable with covariance $\delta\Sigma_e$. (Note that, because of the way the noise is defined in \ref{differential eq} it is not $\delta^2$ but $\delta$). Expanding $q\left(v(r,t_k)\right)$ we have
\begin{equation}\label{discrete time model}
	v\left(r,t_k + \delta \right) =  v(r,t_k) + \delta \alpha \left( -v\left( r,t_k \right) + \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t_k \right)} \right)dr'} \right) + e'(r,t_k)
\end{equation}
which can be simplified to
\begin{equation}\label{discrete time model2}
	v\left(r,t_k + \delta \right) =  (1 - \delta \alpha) v(r,t_k) + \delta \alpha \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t_k \right)} \right)dr'} + e'(r,t_k)
\end{equation}
Now we define the spatial aspect of the model on a regular square $i,j$ grid of neural masses, where the spatial step size $\Delta \mathbf{r}_i = \Delta \mathbf{r}_j = \Delta $ giving
\begin{equation}\label{discrete space}
	v\left(r_{ij},t_k + \delta \right) =  (1 - \delta \alpha)v(r_{ij},t_k) + \delta \alpha \Delta^2 \sum_{i=-I}^{i=I}{\sum_{j=-J}^{j=J}{\tilde{w}\left( r_{ij},r_{ij}' \right)f\left( v\left( r_{ij}',t_k \right) \right)} }+ e'(r_{i,j},t_k).
\end{equation}
where $\tilde{w}(r_{ij},r_{ij}')$ discretized connectivity kernel. In Eq~\ref{discrete space} we have used periodic boundary conditions to alleviate problems associated with edge effects. The connectivity kernel is in the form of a modified Mexican hat, and is described by the sum of three weighted Gaussian functions as
\begin{equation}
	\tilde{w}(r_{ij},r_{ij}') = \mathbf{\theta}^T e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\mathbf{\sigma})^2} + \frac{\left()\left(r_j'-r_j\right)\Delta\right)^2}{(2\mathbf{\sigma})^2}}
\end{equation}
where
\begin{equation}
 \mathbf{\theta}=\begin{bmatrix} \mathbf{\theta}_{1}\\\mathbf{\theta}_{2}\\\mathbf \theta_{3}\end{bmatrix},\quad \mathbf{\sigma}=\begin{bmatrix}\mathbf{\sigma}_1\\\mathbf{\sigma}_2\\ \mathbf{\sigma}_3\end{bmatrix}.
\end{equation}.
The first Gaussian function (parametrized by $\mathbf{\theta}_1, \mathbf{\sigma}_1$) describes local excitation, the second describes lateral inhibition and the third describes medium range excitation.

To generate data will approximate either local field potential recordings or intracranial EEG we have used the observation equation
\begin{equation}\label{ObservationEquation}
    y(r_n,t_k') = \int_\Omega  m( r_n,r' ) v( r',t_k' ) dr'  + \varepsilon ( r_n,t_k ),
\end{equation}
where $\varepsilon( n )\sim\mathcal{N}(0,\Sigma)$, $n$ indexes the iEEG electrodes, and $m(r_n,r)$ is the observation kernel relating the neural field to measurements accounting for electrodes characteristics (spatial averaging). In Eq.~\ref{ObservationEquation} we have indexed time by $t_k'$, which is a decimated version of $t_k$.

All the model parameters are described in Table 1.

\begin{tabular}{ccc}
\hline
Parameter & Value & Reference\\
\hline
$\alpha$ & $100$ s$^{-1}$ &\\
\hline
$\nu_0$ & $1$ s$^{-1}$ &\\
\hline
$v_0$ & $6$ mV &\\
\hline
$\beta$ & $0.56$ spikes/mV$^{-1}$ & \\
\hline
$\delta$ & $1\times10^{-4}$ s &\\
\hline
$\Delta$ & $1$ &\\
\hline
$\theta_1$ &  &\\
\hline
$\theta_2$ &  &\\
\hline
$\theta_3$ &  &\\
\hline
$\sigma_1$ &  &\\
\hline
$\sigma_2$ &  &\\
\hline
$\sigma_3$ &  &\\
\hline
\end{tabular}

\section{Reduced Model For Estimation}

In order to implement standard estimation techniques we propose a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as sequential Monte-Carlo techniques like particle filtering. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
{v_t}\left( {\bf{r}} \right) = {\phi ^T}\left( {\bf{r}} \right){x_t}.
\end{equation}
where $\phi(r)$ is the Gaussian basis functions that are scaled by the the state vector. The connectivity kernel can all be decomposed into a set of basis functions
\begin{equation}\label{DefKernelDecomp}
k( \mathbf{r} , \mathbf{r'} ) = \theta^T\psi(\mathbf{r} , \mathbf{r'} )
\end{equation}
Will will assume we know the parametric form of the basis functions, where the parameter $\theta$ is unknown. Each connectivity basis function can be considered a layer in the WC model, representing short range excitation, surround inhibition and long range excitation. Making substitutions we get
\begin{equation}\label{DecompModel}
{\phi ^T}\left( \mathbf{r} \right){x_{t+1}} = T_s\theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\mathbf{r'}}} \right)f\left( {{\phi ^T}\left( {{\mathbf{r'}}} \right){x_t}} \right)d{\mathbf{r'}}}  - \lambda\phi^T(\mathbf{r})x_t + T_s{e_t}\left( {\mathbf{r}} \right).
\end{equation}
To isolate the state we cross multiply by $\phi(\mathbf{r})$ and integrate over $\mathbf{r}$
\begin{equation}\label{Poo}
\int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}{x_{t+1}}}  = T_s\int_\Omega  {\phi ( \mathbf{r} )\theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  - \lambda\int_\Omega \phi(\mathbf{r})\phi^T(\mathbf{r})d\mathbf{r} x_t  + T_s\int_\Omega  {\phi \left( {\mathbf{r}} \right)e_t\left( {\mathbf{r}} \right)d{\mathbf{r}}}.
\end{equation}
Now we define
\begin{equation}\label{DefGamma}
\Gamma  = \int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
Substituting equation~\ref{DefGamma} into \ref{Poo} and rearranging gives
\begin{equation}\label{SS Model}
{x_{t+1}} = T_s{\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)\theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}} - \lambda x_t + T_s{\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)e_t\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
We can rearrange to get
\begin{equation}\label{Homogeneous SS Model}
	x_{t+1} = T_s\Gamma ^{ - 1}\int_\Omega  {\phi \left( {\bf{r}} \right)\int_\Omega  {f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)\psi^T \left( {{\bf{r}} - {\bf{r'}}} \right)d{\bf{r'}}} d{\bf{r}}}\theta - \lambda x_t + T_s{\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)e_t\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
The disturbance term can be rewritten as 
\begin{equation}
	\Gamma ^{ - 1}\int_\Omega  {\phi \left( \bf{r} \right)e_t\left( \bf{r} \right)d\bf{r}} = \int_\Omega  {\phi \left( \bf{r} \right)\tilde{e}_t\left( \bf{r} \right)d\bf{r}},
\end{equation}
where $\tilde{e}_t(r)~N(0,R'R)$ and $R'$ is the Cholesky decomposition of $\Gamma^{-1}$. This provides the final state-space model
\begin{equation}\label{AbbrevSSModel}
x_{t + 1} = T_sq(x_t)\theta -\lambda x_t + T_s\int_\Omega  {\phi ( \mathbf{r} )\tilde{e}_t( \mathbf{r} )d\mathbf{r}}
\end{equation}
with observation equation
\begin{equation}\label{ObservationEquation}
    y_t( k ) = \int_\Omega  m( k,\mathbf{r} )\phi^T( \mathbf{r} ) d\mathbf{r}x_t  + \varepsilon ( k ),
\end{equation}
